---
title: "Bike share data"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(skimr)
library(summarytools)
library(lubridate)
```

getting the bike share data connected to r
```{r}
dfa = read_excel("DS_bike_share_data.xlsx")
```

Filtering the data to remove bad values of duration

```{r}
dfb <- dfa %>%
  filter(Duration > 0, Duration< 1440)
```

creating histograms of the data

```{r}
hist(dfb$Duration, breaks=120)
```
using facet_wrap to create a histogram for each day of the week

```{r}
dfb %>% ggplot(aes(x=Duration)) + geom_histogram(bins=100, orientation = "x") +
  facet_wrap(~Day_of_week)
```
Creating a scatter plot

```{r}
dfb %>% ggplot(aes(x= hour , y = Duration)) + geom_point()
```
Box plot

How to plot a box and whiskers
```{r}
dfb %>% ggplot(aes( x = Duration, y = hour, group= hour)) + geom_boxplot()
```

How many bikes are being used at any given time of the day.

### how many riders
```{r}
dfe=dfb %>% filter(mday(started_at)==3) %>% slice_sample(n=100)
```

select the columns we need 
```{r}
dfe %>% select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end) 
```
pivot longer so we can work with the data.
```{r}
dfe %>% select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t)
```
createa  graph with the new longer data set.

```{r}
dfb %>% 
  select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  mutate(riders=cumsum(increment)) %>% 
  ggplot(aes(t,riders)) + geom_step() 
```
create a small sample to show the general shape of the graph.


```{r}
dfe %>% select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  mutate(riders=cumsum(increment)) %>% 
  ggplot(aes(t,riders)) + geom_step()
```
creating a facet wrap graph to display each day of the month


```{r}
dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
  select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  mutate(riders=cumsum(increment)) %>% 
  ggplot(aes(t,riders)) + geom_step() +
  facet_wrap(~mday(t),scales = "free_x",ncol = 7)
```

Pivot longer

one column the type and the other the "t"

mutate to create an "increment" column ( +1 or -1)
use case when

arrange based on time 
mutate bikes using cumsum

```{r}
dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
 select(rideable_type,start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  group_by(rideable_type) %>%
  mutate(riders=cumsum(increment)) %>% filter(mday(t)<=7) %>%
    ggplot(aes(t,riders,color=rideable_type)) + geom_step() +
  facet_wrap(~mday(t),scales = "free_x",ncol = 7)
```
```{r}
dfr = dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
 select(rideable_type,start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  group_by(rideable_type) %>%
  mutate(riders=cumsum(increment))
```

Creating dfr1 to simplify the data

```{r}
dfr1 = dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
 select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  group_by() %>%
  mutate(riders=cumsum(increment))
```

How many riders in an hour time interval. ( max)

```{r}
dfrh = dfr1 %>% group_by (hour(t)) %>%
 summarise( max_riders = max(riders), ave=mean(riders))
day
```
Creating a count of how many riders start and finish at each station.
```{r}
 dff=dfb %>%
  group_by(start_station_id) %>%
  summarise(amount_of_riders = n())
```

Creating the end station data

```{r}
dfh=dfb%>%
    group_by(end_station_id) %>%
  summarise(amount_of_riders = n())
```

Joining the data

```{r}

```


