---
title: "Class enrollment Data:MA132"
author: "Richard Pandel"
date: "9/21/2020"
output: html_notebook
---

The intention of this analysis is to accurately predict the amount of students who enroll into MA 132 next semester.



prepare workplace

## Load packages

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(skimr)
library(summarytools)
```
```{r}
dfa= read_excel("CU_ACAD_DEPT_ENRL.XLSX") %>%
  clean_names( )
```
```{r}
skim(dfa)
```

### Tidy Data

Let's separate column term2 into to variables

```{r}
dfb = dfa %>% separate (term2, into=c("semester","year")," ")
```
### clean and subset the data

sd get rid of the following variables:

*acad_org
*current_status
*Instructor
*course_id

Filtering to only include:

*math courses
*include catalog numbers for MA131 and 132
```{r}
dfc = dfb %>% select(-acad_org,-current_status,-instructor,-course_id) %>% 
  filter(subject =="MA")
  filter(catalog == ("131","132"))
```
Filtering to remove duplicates
```{r}
dfc<-subset(dfb, component!="DIS")
```
```{r}

```

